Type Error Telemetry at Scale

### 2023-09-24 programming23/paper.tex
split type err / forced strict by mode

\subsection{Forced Strict Errors by Mode}

FILL

How do the 72 million forced strict divide across modes?

What about type errors, same distribution?!

@@

To deal with pathologies such as the worst-case time for ML-style type
inference~\cite{m-popl-1990,ktu-caap-1990}, the typechecker
has internal limits that restrict the problems it will attempt to solve.
Hitting a limit triggers one of the following errors:
\code{CodeTooComplex},
\code{NormalizationTooComplex}, or
\code{UnificationTooComplex}.
A casual user should never see these errors.\todo{IDEALLY never see?
  They CAN see them, right?}

>>>>

\subsection{Errors by Mode}

\begin{figure}[t]
  %% strict
  %%  te total 22771
  %%  mean 6.9 std 26.24592591869705 med 1 max 385
  %% 
  %%  fs total 114259
  %%  mean 34.9 std 155.41955823164017 med 1 max 4789
  %% 
  %% nonstrict
  %%  te total 362964
  %%  mean 2.3 std 66.44456626891912 med 0 max 13511
  %% 
  %%  fs total 7546175
  %%  mean 48.6 std 270.60578740369266 med 3 max 16712

  \begin{tabular}{l}
    600K type errors         = \pct{35} \mnocheck{}, \pct{61} \mnonstrict{}, \pct{4} \mstrict{}
    \\
    72M forced scrict errors = \pct{89} \mnocheck{}, \pct{10} \mnonstrict{}, \pct{1} \mstrict{}
    %% really 0.16 strict, very low!!!
  \end{tabular}
  \caption{}
  \label{f:error-by-mode}
\end{figure}

\Cref{f:error-by-mode} shows how the type errors and forced strict errors
in the data are split across type analysis modes.
\FILL{} add averages.

Most type errors come from \mnonstrict{} mode, despite it having an order
of magnitude fewer users that \mnocheck{}.
This is unsurprising because \mnocheck{} reports only syntax errors,
whereas \mnonstrict{} reports several others kinds~(detailed in the next section).
Finally, \mstrict{} mode accounts for a small fraction of type errors, despite
being more picky, because so few creators use it.

Forced strict analysis applies the same typechecker to every module, regardless
of mode, in the hope of enabling IDE tools.
Consequently, the breakdown of forced strict errors matches the overall usage
data~(\cref{f:dataset-overview}): almost \pct{90} for \mnocheck{},
\pct{10} for \mnonstrict{}, and \pct{1} for \mstrict{}.
Type analysis finds a large number of errors in \mnocheck{} code;
unfortunately, we have no insight into whether those errors are true
positives or whether the analysis led to useful IDE suggestions.

@@

To deal with pathologies such as the worst-case time for ML-style type
inference~\cite{m-popl-1990,ktu-caap-1990}, the typechecker
has internal limits that restrict the problems it will attempt to solve.
Hitting a limit triggers one of the following errors:
\code{CodeTooComplex},
\code{NormalizationTooComplex}, or
\code{UnificationTooComplex}.
These errors can appear in Script Analysis just like any other
error, but one design goal of Luau is that code rarely hits the limits.
A typical creator should never see these errors.
